Formalizing Stack Safety as a Security Property

### 2022-11-01 submission5/noexceptions.tex
Moved more detailed explanation into earlier section

@@

Otherwise, we do not assume that the code is reasonable\apt{still vague} or adheres to any particular
calling convention or source construct. In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

@@

This is a strong threat model, but hardware and timing attacks are out of scope,
and our properties are termination insensitive as a result of the enforcement mechanism
(below).

@@

For convenience, we will abbreviate multiple updates of a mapping
\(\mach[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(\mach \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).

@@

A pending activation also carries the addresses to which the program counter
and stack pointer should point in the event of a return. We represent the call
stack with a list of such triples. A {\it context} is a pair of the current activation's view
and a list of pending ones.
The initial context is \(\context_0 = (V_0, [ ~ ])\).
\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]

@@

From a machine state and a context, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with the same operations. Note that \(Op\) takes as its state argument
\(\mach\), the state before the step.

@@

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret in its arguments and initializes a variable to contain
potentially sensitive data.\apt{Arguments to {\tt main} are overly complicated.
  There's no need to match reality (and in fact {\tt argv} is really a {\tt char*}
  anyway).}\apt{As we discussed, it would be simpler to represent ``publish'' by
  writing to a specified global rather than with a call.}
It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will publish the results of {\tt f}.
\apt{Would be better to use symbolic names instead of inetegers for {\tt jal}
  targets. Figure at bottom is good, but also needs additional labelling to
  show addresses relative to the new sp within {\tt f}.}

\begin{figure}
  \begin{subfigure}{\columnwidth}
    {\tt
      void main(int argc, int* argv) \{

      ~ ~ int secret = argv[1];

@@

      ~ ~ ~ ~ publish(secret);

@@

      ~ ~ ~ ~ publish(res);

@@

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a1,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal 100,ra}{\(\mathbf{call} ~ 100 ~ \emplist ~ \emplist\)}
      \labeledrow{20:}{sw r0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
      \labeledrow{52:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.
\apt{In (a,b), {\tt 12(sp)} maps to {\tt sensitive}, not {\tt secret} (I think!).
  It would be nice if (a) didn't have to push a frame (i.e. again by making publication
  just a global write) and if all the examples of {\tt f} were the same length, e.g.
  by adding {\tt nop}s.}

In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
instead return that value so that {\tt main} makes the call to {\tt publish},
as in Figure \ref{subfig:indirect}.

@@

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi sp,sp,-8}{\(\mathbf{alloc} ~ (-8,8)\)}
      \labeledrow{104:}{sd ra,0(sp)}{}
      \labeledrow{108:}{lw a0,12(sp)}{}
      \labeledrow{112:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0} ~ \emplist\}\)}
      \labeledrow{116:}{lw ra,0(sp)}{}
      \labeledrow{120:}{addi sp,sp,8}{\(\mathbf{dealloc} ~ (0,8)\)}
      \labeledrow{124:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} directly}
    \label{subfig:direct}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a0,12(sp)}{}
      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} indirectly}
    \label{subfig:indirect}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{li a5,42}{}
      \labeledrow{104:}{sw a5,4(sp)}{}
      \labeledrow{108:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking {\tt sensitive}}
    \label{subfig:integrity}
  \end{subfigure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi ra,ra,20}{}
      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking control flow}
    \label{subfig:WBCF}
  \end{subfigure}

@@

\apt{This para is a bit of an orphan.}
At the start of execution the program counter is 0 and the stack pointer -4 \apt{or is it 0?},
with the stack growing down. The functions {\tt f} and {\tt publish}
are at addresses 100 and 200, respectively.

\apt{Need a smoother transition here. Be explicit that we are now going to start considering the security
  state.}

Figure \ref{fig:exec1} shows how the program counter, stack pointer, and context
develop over several steps, with labels on the steps that allocate space and then make
a call. \(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\) in this diagram;
events are omitted.\apt{graphics on right of figure need to be keyed to corresponding steps.}

@@

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l | l || l |}
      \hline
      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
      \hline
      1 & 4 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      4 & 16 & -20 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & 100 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[(V_1, 20, -20)]\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      \(n\) & ? & ? & \(V_2,[(V_1, 20, -20)]\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      \(n+1\) & 20 & -20 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

@@

On step \(n\), {\tt f} returns, and we restore the topmost inactive principal,
{\tt main}'s principal. \apt{This $n$ is a bit mysterious. Again, if all the snippets
  were the same length, you could use a concrete number here.}

\apt{Need a bridging sentence here, along the lines of: ``Now we consider how to define desirable
  stack safety properties that rule out various kinds of bad behavior.''}

@@

But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\))\apt{-20?},
as would be the case in attack (d)? Then it has violated the WBCF property.
This property can be phrased in terms of a single step: for every transition
labeled as a return, the program counter and stack pointer at the next state
match those recorded in the top pending activation. The table in Figure \ref{fig:exec1}
assumes that WBCF is followed.

@@

Integrity is more complicated, and requires us to compare the state at the start of
the call to {\tt f} to that on return. Intuitively, the integrity of {\tt main}
is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
to be identical to when it made the call.

In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses -20 through -1 (allocated as private data) and callee-saved registers such as
the stack pointer and return address.
Note that the callee-saved registers may well have changed during the call---but if
so, the callee is obligated to restore them to their prior value.

One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
in a way that is harmless---or that is rendered harmless by the enforcement mechanism!
\apt{This is still confusing, because the ``rendered harmless'' part applies just as much
  to the strict integrity we've been discussing up to now.}
This is the case in a lazy micro-policy\apt{presumes we've described these a bit in the introduction
  (which we should)}: the monitor permits the caller to write into the
callee's frame, but taints that write so that the callee cannot read it later without
triggering a failstop. In such a scenario we will consider a value {\it irrelevant}
if it could be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every
element except those in \(\components\). The elements of \(\components\) are said to be
irrelevant in \(\mach\) if every variant state produces an equivalent trace
(see Section \ref{sec:events}.)

So, to define integrity, we need to know when a caller has been returned to;
this is the first state after the call in which the depth of the call stack
matches the state before the call. Integrity holds if, at that return state,
any element that is \(\sealed\) under the callee's view is either restored
to its original value or is irrelevant.

Returning to our example: at step 4, when {\tt main} makes its call to {\tt f},
the call stack in the context is empty. It makes its call, and at step 5, {\tt main}'s
frame, including {\tt secret} and {\tt sensitive}, is mapped to \(\sealed\) from the
perspective of {\tt f}. So, the next time the call stack is empty (implicitly, when
we have returned to {\tt main}), elements of that frame must either match their
values at step 4, or be irrelevant to future computation.\apt{picture?}

@@

We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
means that its callee's behavior is dependent only on publicly visible data,
not the caller's private state. \apt{Spell out the implications of this for initialization.}
As we saw in the examples, we must consider both the observable events
that the callee produces during the call and the changes that the callee makes to the state that might
affect the caller after the callee returns.

Consider the state \((\mach, (V_2,\sigma))\) at step 5. If we take a variant state over
the set of element(s) that are \(\sealed\) or \(\unsealed\) in \(V_2\),
{\it internal confidentiality} means
that, if we execute both states until they return, they will produce equivalent sequences
of events.

>>>>

\begin{abstract}

``Stack safety'' is a term associated with a
variety of compiler, run-time, and hardware mechanisms for protecting stack
memory. But unlike heap memory, the stack does not correspond to any single
high-level language concept. The result is that
these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of the bad
behaviors that they prevent. Such a specification is especially desirable
for an unusual enforcement mechanism, like the ``lazy'' stack safety
micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.

We propose a formal characterization of ``observable'' stack safety based on
concepts from language-based security. Stack safety is a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and confidentiality properties (``the caller's and callee's behavior are
insensitive to one another's private state,''), both compatible with lazy tagging, as well as
control-flow property.

The stack governs a large number of language features that are often excluded
from discussion of stack safety. Our stack safety property targets a system with
both caller- and callee-saved registers, coroutines, arguments passed on the stack,
exceptions, [tail call elimination, heaps, and concurrency.]

We use these properties to validate Roessler and DeHon's micro-policies
via property-based random testing. Our testing successfully detects violations
in several broken variants, including Roessler and DeHon's original lazy policy.

\end{abstract}

\newcommand{\paragraphx}[1]{\emph{#1.}}

\section{Introduction}

The call stack is a perennial target for low-level attacks, leading to a
range of dire consequences, from leakage or corruption of private stack data
to control-flow hijacking. To foil such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
  \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi

The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more abstract characterization to the reader's intuition.  But these
mechanisms can be quite intricate, and the behaviors they aim to prevent are
subtle and varied. It can be hard to guess, based only on informal
intuitions, whether a given mechanism actually blocks all potential
attacks---or, conversely, whether it is overly conservative and disallows
system behaviors that are actually safe.
To settle such questions with confidence,
we need a precise, generic, and formal specification for stack
safety, as a basis both for comparing the security claims of different
enforcement techniques and for validating that these claims
are met by particular implementations. With the notable exception of the model used
in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
the literature.

We propose a novel characterization of stack safety using the tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into
the {\em integrity} and {\em confidentiality} of the caller’s local state
during the callee's execution, and the confidentiality of the calle's local state
afterward, as well as the control flow protection of {\em well-bracketed control
  flow (WBCF)}~\cite{SkorstengaardSTKJFP}. We extend WBCF to a setting with exceptions,
in which there may be multiple points at which a callee can ``return.''

A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.).
Caller confidentiality is especially interesting. It is based on a traditional
notion of noninterference, but whereas ordinary noninterference
is an end-to-end hyper-property on whole program runs, caller
confidentiality is a nested form of noninterference that applies to
part of a run, requiring that the callee’s behavior is
invariant under hypothetical scrambling of the caller’s stack
but allowing the caller to access its own data upon return.

Our threat model is very strong, allowing the attacker
to execute arbitrary code, including attempting to smash
the stack to disrupt the program’s control flow. Our stack safety
properties demand that, even in the presence of such attackers, confidentiality
and integrity still apply.
%

For confidentiality and integrity, our properties are {\em observational},
meaning that they allow a function activation to read from and write to other activations'
data, as long as such reads and writes do not affect the observable behavior of the
system. This permissiveness enables our properties to handle lazy enforcement mechanisms.

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification.  We
use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and DeHon's micro-policies correctly abort the ones that
attempt to violate one of our properties. Furthermore, we
%
check that the testing framework is able to generate counterexamples
that violate our properties but are \emph{not} halted by incorrect
enforcement variants---both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself.

We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
which memory within each stack frame is tagged with the depth of
the function activation that owns the frame and access is
permitted only when an activation at that depth is currently executing, validates our
stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%

%For ease of exposition, we initially assume a single simple
%stack with no sharing between callers and callees: all parameters and return
%values are passed in registers.  Later, we will introduce enhanced versions
%supporting
%(1) passing of scalar stack data,
%(2) callee-saves registers calling conventions, and
%(3) a coroutine system with a static layout.

In sum, we offer the following contributions:

\begin{itemize}
\item We formalize permissive stack confidentiality and integrity and integrity properties.
  Both are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in the system behavior %(\cref{sec:lazy}).
\item We apply these definitions to a complex system with features such as argument passing
  on the stack, callee-saves registers, exceptions, and a simple coroutine system% (\cref{sec:ext}).
\item We use property-based random testing to validate the relationship between
  our properties and micro-policies %(\cref{sec:testing})
\end{itemize}
%\Cref{sec:relwork,sec:future} discuss related and future work.

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as the testing implementation.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

@@

This is a strong threat model, but hardware and timing attacks are out of scope,
and our properties are termination insensitive as a result of the enforcement mechanism.

@@

@@

For convenience, we will abbreviate multiple updates of a mapping
\(V[\component_0 \mapsto sc_0][\component_1 \mapsto sc_1]\dots\)
as \(V \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).

We represent the call stack with a list of views representing pending inactive
principals, ranged over by \(\sigma\). A {\it context} is a pair of the current
activation's view and that list. The initial context is \(\context_0 = (V_0, \emplist)\).
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]

@@

From a machine state and a context, we create a {\it combined state}
\(s = (\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with the same operations. Note that \(Op\) takes as its state argument
\(\mach\), the state before the step.

@@

\subsection{Policy State}

In addition to its elements, a system may contain additional ``policy'' state that carries
information relevant to a hardware enforcement mechanism. We assume that policy state
does not influence execution, except to cause the system to failstop, which is modeled
as a state stepping to itself. A failstop is therefore indistinguishable from silent
divergence.

Formally, we can treat our machine state as
\(\mach = (\mach[\cdot],\langle) \mach \rangle \in \MACHS
::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
Then \(\langle \mach \rangle\) is the security state. We assume that if
\(\mach[\component] = \nach[\component]\) for all \(\component\),
\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
then \(\bar{\psi_1} = \bar{\psi_2}\) and either
\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
for all \(\component\).

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We will assume for this system that events are stores
to a special address, called {\tt out}.

{\tt main} then calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will publish the results of {\tt f}.

\begin{figure}
  \begin{subfigure}{\columnwidth}
    {\tt
      void main(int secret) \{

@@

      ~ ~ ~ ~ *out = secret;

@@

      ~ ~ ~ ~ *out = res;

@@

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a1,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
      \labeledrow{20:}{sw r0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{sw a0,out}{}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
      \labeledrow{52:}{sw a0,out}{}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.

In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in Figure \ref{subfig:indirect}.

@@

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{sw a4,out}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} directly}
    \label{subfig:direct}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{mov a4,a0}{}
      \labeledrow{108:}{sw zero,-4(sp)}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} indirectly}
    \label{subfig:indirect}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{li a5,42}{}
      \labeledrow{104:}{sw a5,4(sp)}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking {\tt sensitive}}
    \label{subfig:integrity}
  \end{subfigure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi ra,ra,20}{}
      \labeledrow{104:}{nop}{}
      \labeledrow{108:}{nop}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking control flow}
    \label{subfig:WBCF}
  \end{subfigure}

@@

Figure \ref{fig:exec1} shows how the security context
develops over several steps, with labels on the steps that allocate space and then make
a call. Both program counter and stack pointer are initially 0, and the stack pointer
is shown for convenience.
\(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\)
in this diagram; events are omitted.

@@

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l || l |}
      \hline
      Step & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
      \hline
      1 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\vdots\)} \\
      \hline
      4 & -20 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(\vdots\)} \\
      \hline
      8 & -20 & \(V_2,[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      9 & -20 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

@@

On step 8, {\tt f} returns, and we restore the topmost inactive principal,
{\tt main}'s principal. Now we will discuss the intuition behind how this
structure translates into stack safety properties that rule out different
classes of bad behavior.

@@

But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
\(\SP \neq -20\)), as would be the case in attack (d)? We consider this to
violate well-bracketed control flow (WBCF). WBCF is a relationship between
call steps and their corresponding return steps: after the return, the program
counter should be at the instruction following the state before the call,
and the stack pointer should be the same as it was before the call.

Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
20 to the return address and then returns, thus bypassing the check and outputting
{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 20,
so we define a predicate on states \(ret\) that holds when the
program counter and stack pointer are both 20.

Then we identify the state immediately following the return, if one occurs ---
this will be the first state in which the pending call stack is smaller than it was
inside of the call.
WBCF requires that if \(s\) after that return, \(ret ~ s\) will hold.
For nested calls, where the pending stack is initial larger, the same principal
applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.

To demonstrate why WBCF also cares about the stack pointer, consider an attack which
returns with the stack pointer at -12 instead of -20. Now, compare the layout of the
stack relative to the stack pointer as it should be (\(\SP\)) to the one that results by
increasing the stack pointer before return (\(\SP'\)):

\begin{tabular}{| l | l | l | l | l |}
  \multicolumn{1}{r}{\(\SP \downarrow\)} &
  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
  \hline
  res & sens & sec & ra & ra \\
  \hline
\end{tabular}

In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
read part of the return address, and then it will attempt to output
{\tt res}, but instead output {\tt secret}! Even absent other
kinds of data protection, the stack pointer {\it must} be restored
for the program to behave predictably.

@@

Stack integrity applies a similar notion to all of the memory that a function has
allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
upon allocation, the view labeled is \(\object\), and when a call is made, it instead
becomes \(\sealed\). In more sophisticated systems discussed later, we will distinguish
sealed objects from those that are intentionally passed to the callee.
Intuitively, the integrity of {\tt main}
is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
to be identical to when it made the call.

In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses -20 through -1 (allocated as private data) and callee-saved registers such as
the stack pointer and return address. Note that callee-saved registers often change
during the call---but if so, the callee is obligated to restore them to their prior value.

One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
in a way that is harmless. Or, an enforcement mechanism might defer the checking
for illegal writes until later in the execution, entering an error state at some
point after return but before the illegal write is able to damage the system.
This is the case in a lazy micro-policy: the monitor permits the caller to write into the
callee's frame, but taints that write so that the callee cannot read it later without
triggering a failstop. In such a scenario we will consider a value {\it irrelevant}
if it could be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\)
are {\em \(\components\)-variants} if they agree on the value of every
element except those in \(\components\), and share the same policy state.
The elements of \(\components\) are said to be
irrelevant in \(\mach\) if every variant state produces an equivalent trace
(see Section \ref{sec:events}.)

So, to define integrity, we again need to know when a caller has been returned to,
and we use the same mechanism of checking the depth of the call stack.
Integrity holds if, at that returned state,
any element that is \(\sealed\) under the callee's view is either restored
to its original value or is irrelevant.

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation}
  \label{fig:variant}
\end{figure}

The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). Figure \ref{fig:variant} shows the state at step 5, just after the call,
assuming that {\tt sec} is 5. Similar to WBCF, we will define
\(int ~ s\) as a predicate on states that holds if, of the sealed
addresses in state 5, all values that differ in \(s\) are irrelevant. We once again
require that \(int\) hold on the state following the matching return.

That state is reached at step 9. Since {\tt sensitive} has changed, we must determine whether
it is irrelevant. We consider a variant state in which it has any other value, arbitrarily
choosing 43. As execution continues after the return from the original state, it
passes the check on {\tt sensitive}, while the execution from the variant does not, resulting
in differing outputs. Therefore {\tt sensitive} is not irrelevant, so \(int\) does not hold,
and integrity has been violated.

@@

We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
means that its callee's behavior is dependent only on publicly visible data,
not the caller's private state. This also requires that the callee initialize
memory before reading it.
As we saw in the examples, we must consider both the observable events
that the callee produces during the call and the changes that the callee makes to the state that might
affect the caller after the callee returns.

Consider again the state \((\mach, (V_2,\sigma))\) at step 5, this time with the attacker from
Figure \ref{subfig:direct}. We take a variant state over the set of element(s) that are
\(\sealed\) or \(\unsealed\) in \(V_2\), which happens to be the entire stack (see
Figure \ref{fig:variant2}.) If we take a trace of execution from each state until it returns,
the traces may differ, in this case outputting 5 (the original value of {\tt sec}) and
4 (its value in the variant) respectively. This is a violation of
{\it internal confidentiality}.

\begin{figure}
  \includegraphics[width=\columnwidth]{variants2.png}
  \caption{Internal Confidentiality Violation}
  \label{fig:variant2}
\end{figure}

But, in Figure \ref{subfig:indirect}, we also saw an attacker that exfiltrated the secret
by reading it and then returning it, in a context where the caller would output the returned
value. Figure \ref{subfig:variant3} shows the behavior of the same variants under this attacker,
but in this case, there is no output during the call. Instead the value of {\tt secret} is
extracted and placed in {\tt a0}, the return value register. To illustrate the subtlety
of identifying a leak, {\tt f} has also stored a 0 below the stack pointer.

\begin{figure}
  \includegraphics[width=\columnwidth]{variants3.png}
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
